---
title: "Mini Intro to Purrr package"
author: "Camila Vargas"
date: "3/29/2022"
output: html_document
---
This is a micro-introduction to the `purrr` package. The intention of this script is to show some of the concepts and functions related to `purrr` to start a conversation about `purrr`, for loops and apply functions at the NCEAS data-science-chat.

## Introduction

_What is `purrr`?_
- Part of the `tidyverse`
- Takes the tasks that the `apply` suit of functions do in base R.

_Why `purrr`?_
- To apply operations across many dimensions of your data
-  improving your ability to keep even complex analyses "tidy" (enhancing functional programming)


**Goal of this chat:**
- Compare three ways of doing the same operations: loops, apply and `purrr`
- Get you excited about `purrr` and its power
- Learn something new


When reading about `purrr`, two concepts that immediately come up **Lists** and **Functional Programming**.

**Functional Programming** - trying to use functions (preferably "pure" ones) to accomplish most of your complex and repetitive tasks (don't copy and paste more then 3 times - H. Wickham)

Functions in R take any number of named inputs, do operations on them, and returns the last thing produced inside the function (or whatever you specify using return)

**Lists** are powerful objects that allow you to store all kinds of information in one place (not only atomic vectors, vectors with different length).
They can also be a pain to deal with, since we are no longer in the nice 2-D structure of a traditional data frame.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(repurrrsive)
library(listviewer)
```


## Loops
Main use of loops it to iterate in R in various scenarios.

The long repetitive way of doing an operation looks like
```{r}
# Create a vector of the mean values of all the columns of the mtcars dataset
mean_vec <- c(mean(mtcars$mpg), mean(mtcars$cyl), mean(mtcars$disp), mean(mtcars$hp),
              mean(mtcars$drat), mean(mtcars$wt), mean(mtcars$qsec), mean(mtcars$vs),
              mean(mtcars$am), mean(mtcars$gear), mean(mtcars$carb))

mean_vec

```

The loop way

```{r}
# Create an "empty" vector
mean_vec_loop <- vector("double", ncol(mtcars))

# Populate the vector with means
for (i in seq_along(mtcars)) {
  mean_vec_loop[[i]] <- mean(mtcars[[i]])
}

mean_vec_loop
```

- Results are the same
- Not much differences in computing time
- But, no-loop long code - prone to errors


## `apply`
Higher Order Functions. Similar to what we did earlier, these are functions that can take other functions as an argument. 

The benefit of using higher order functions instead of a for loop is that they allow us to think about what code we are executing at a higher level. Think of it as: "apply this to that" rather than "take the first item, do this, take the next item, do this..."

```{r}
# Create a list/vector of the mean values of all the columns of the mtcars dataset
lapply(mtcars, mean) %>% head # Returns a list
sapply(mtcars, mean) %>% head # Returns a vector
```

- Main argument to use `apply` or `purrr` is not execution time is simplicity and readability of the code.

## `purrr`
Main function in this package is `map()`. This function is used in a similar way that `apply`.

```{r}
# Create a list/vector of the mean values of all the columns of the mtcars dataset
map(mtcars, mean) %>% head # Returns a list

map_dbl(mtcars, mean) %>% head # Returns a vector - of class double
```

Other `purrr` functions

`map_lgl()` - makes a logical vector.
`map_int()` - makes an integer vector.
`map_dbl()` - makes a double vector.
`map_chr()` - makes a character vector.

## More `purrr` examples

Using the Game of Throne data from the `repurrrsive` package. Where each list has information of one character of the show.

```{r}
str(got_chars, list.len =  3)
```

Cool interactive approach to view lists using the `listviewer` package.
```{r}
listviewer::jsonedit(got_chars)
```


So, how do we start poking around in this database?

Suppose we wanted only the first 5 characters in the list

```{r}
got_chars[1:5] %>% 
  str(max.level = 1)
```

Now, suppose that we jut want to look at the name of the first 5 characters. 

You might think that `got_chars[[1:5]]$name` would do the trick...

```{r, error = T}
got_chars[[1:5]]$name
```

Nope.

So we could do

```{r}
names <- vector(mode = "character",5)
for (i in 1:5){
  
  names[i] <- got_chars[[i]]$name
}
names
```

That works, but certainly not ideal.

Enter `purrr`

```{r}
got_chars[1:5] %>%
  map_chr('name') #goes through the character list form 1 to f and prints out the name
```

Using `listviewer`

```{r}

got_chars[1:5] %>%
  set_names(map_chr(.,'name')) %>%
  listviewer::jsonedit()

```



## Resources
Examples and information comes from the sources below.

- [Dan Ovando EcoDataScience purrr workshop](https://github.com/camilavargasp/learn-purrr)
- [To purrr or not to purrr by Nicolas Attalides](https://www.r-bloggers.com/2018/05/to-purrr-or-not-to-purrr/)
